# Matrices in R
**Created by John Palowitch, UNC Chapel Hill**

**Edited by Robin Cunningham, UNC Chapel Hill**

#Amanda Liu PID:730042603

### Conceptual Questions 
Please read the document ***Matrices in R.docx*** before completing this assignment. This assignment has two goals:

-Learning new commands in R from a written source.

-Using R for Matrix Calculations
<br><br>
1. Suppose:
\[
\mathbf{A} = \begin{bmatrix}2&3&4&5\\-1&0&3&2\end{bmatrix},\;\;\;\;\mathbf{B} = \begin{bmatrix}4&-1&10&0\\-2&-2&5&6\end{bmatrix}
\]
\[
\mathbf{a} = \begin{bmatrix}6&2&-11&1\end{bmatrix},\;\;\;\;\mathbf{b} = \begin{bmatrix}4\\2\\3\\0\end{bmatrix},\;\;\;\;\mathbf{c} = \begin{bmatrix}-3\\-10\end{bmatrix}
\]
    a. Write down the dimensions of the following (you can use a comment box that you create). If the operation is not possible, write "N.P."
        i. $\mathbf{A}^T\mathbf{B}$   4X4 (first one done for you!)
        ii. $\mathbf{B}\mathbf{A}^T$
        iii. $\mathbf{A}\mathbf{B}$
        iv. $\mathbf{A}^T\mathbf{A}$
        v. $\mathbf{b}^T\mathbf{A}$
        vi. $\mathbf{a}\mathbf{A}^T$
        vii. $\mathbf{B}\mathbf{b}$
        viii. $\mathbf{B}\mathbf{c}^T$
        ix. $\mathbf{a}\mathbf{c}$
        x. $\mathbf{b}\mathbf{a}$
```
1. 4X4
2. 2X2
3. N.P
4. 4X4
5. N.P
6. 1X2
7. 2X1
8. N.P
9. N.P
10.4X4
```

    b. Again, write down the dimensions of the following (on this sheet, next to the expression). If the operation is not possible, write "N.P."
        i. $(\mathbf{A}^T\mathbf{B})^{-1}\mathbf{a}$
        ii. $\mathbf{c}(\mathbf{B}\mathbf{A}^T)^{-1}$
        iii. $(\mathbf{B}^T\mathbf{B})^{-1}\mathbf{B}^T\mathbf{c}$
        iv. $\mathbf{b}^T\mathbf{A}^T\mathbf{c}$
        v. $(\mathbf{B}\mathbf{B}^T)^{-1}\mathbf{B}\mathbf{b}$
```        
1.N.P
2.N.P
3.4X1
4.1X1
5.2X1
```

### Applied Questions 
2. In a code block below, create and print the following matrices and vectors:
\[
\mathbf{A} = \begin{bmatrix}-6&0&1\\1&10&7.5\end{bmatrix},\;\;\;\;\mathbf{B} = \begin{bmatrix}-15&1&1\\5.25&1&2.5\end{bmatrix}
\]
\[
\mathbf{a} = \begin{bmatrix}6&2&1\end{bmatrix},\;\;\;\;\mathbf{b} = \begin{bmatrix}2\\-7\end{bmatrix}
\]
```{r}
A = matrix(c(-6,0,1,1,10,7.5), nrow = 2, byrow = TRUE)
B = matrix(c(-15,1,1,5.25,1,2.5), nrow = 2, byrow = TRUE)
a = matrix(c(6,2,1),nrow = 1,byrow = TRUE)
b = matrix(c(2,-7),nrow = 2,byrow = TRUE)
A
B
a
b
```



3. Compute (with **R**, in a new code block below) and print the following products:
    a. $\mathbf{A}^T\mathbf{B}$
```{r}
(t(A)%*%B)

```

    b. $\mathbf{B}\mathbf{a}^T$
    
```{r}
(B%*%t(a))
```
    c. $\mathbf{b}^T\mathbf{A}$
```{r}
(t(b)%*%A)
```

    d. $\mathbf{A}^T\mathbf{b}$
```{r}
(t(A)%*%b)
```
    e. $\mathbf{a}^T\mathbf{a}$
```{r}
(t(a)%*%a)
``` 

    f. $\mathbf{a}\mathbf{a}^T$
```{r}
(a%*%t(a))
```
    g. $(\mathbf{A}^T\mathbf{A})^{-1}$
```
N.P
```
    h. $(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{b}$
```
N.P
```
    i. $\mathbf{b}^T\mathbf{A}\mathbf{a}^T$
```{r}
(t(b)%*%A%*%t(a))
```
    j. $\mathbf{a}\mathbf{B}\mathbf{A}^T\mathbf{b}$
```
N.P
```



