---
title: 'Case Study 1: College Basketball'
author: Yihan Qiu(730051944) & Amanda Liu (730042603)
output: html_document
---

In this Case Study, you will refresh your memory of STOR 155 while you learn some basic commands and tools for analyzing data with **R**.  We'll be looking at some data from college basketball games last year.

Run the following **R** code to load the data into your RStudio and take a look at it.

## Summarizing data

```{r, eval = TRUE}
# Load dataset
bball = read.csv("http://kbodwin.web.unc.edu/files/2016/06/basketball.csv")

# Look at dataset
head(bball)
summary(bball)
ncol(bball)
nrow(bball)
tail(bball)

```

The command `read.csv( )` will read a dataset into R from your computer or from online.  "csv" stands for "comma separated value", a common file type where the data is listed in a text file, with variables separated by commas.  For now, you don't need to worry about the details of `read.csv( )`.  Once you have loaded the data, the command `summary( )` will tell you about the variables in the dataset and their values.  Another useful function is `head( )`, which shows you the first 6 rows of the dataset.

***

### Question 1:  

a) Look at the outputs of `summary(bball)` and `head(bball)`, and describe the variables using vocabulary from STOR 155.  
```
Team,team location, opponent and team result are qualitative/categorical variable;X,date are team score and opponent score are numerical, especially discrete variable.
```

b) If `head( )` shows the first 6 rows of the dataset, what command do you think might show the *last* 6 rows?  Try out your proposed function and see what happens
``` 
tail();
```

c) Try the commands `ncol( )` and `nrow( )`.  What do these do?  How could you get the same information from `head( )`, `summary( )`, and/or the command you figured out in part (b)?
```
ncol()is the number of colum and nrow() is the number of row of the dataset.ncol() could be simply counted in the head() and summary() datasets. nrow() can be figured out by commands tail(bball) and looking at the index of the last event.
```

***

Sometimes, we will want to look at individual entries, rows, or columns of our data matrix.  We can do this using brackets `[ ]` after our dataset.  We can also look at a variables (columns) by name using the *$* symbol.  Try the following examples.

```{r, eval = TRUE}
# Look at a single row
bball[123, ]

# Look at a single column
bball[ , 4]
bball$Team.Score

# Look at a single entry
bball[123, 4]
bball$Team.Score[123]

# Calculate mean, median, variance, and standard deviation
mean(bball$Team.Score)
median(bball$Team.Score)
var(bball$Team.Score)
sd(bball$Team.Score)

```

***

### Question 2:

a) What is the difference between `mean(bball$Team.Score)` and `mean(bball[,4])`?  Why might it be useful to have two ways to get access the variable `Team.Score`?
```
The first one access through the name of colum and the other one access through the number of colum. It is more flexible to have two ways to access the variable and we can chose the way that is more convenient for us.(Like when the name is hard to type and when there are too many colums to count)
```

b) In plain English, what were the events of the game represented by the first row of the dataset?
``` 
Old Dominion scored 67 in the game versus Niagara which scored 50 and won in Nov,13,2015. The game was held by a neutral location where neither of the team is the home team.
```
*(Note:  If you don't know much about basketball - for example, if you don't know what it means to play a game "Home" versus "Away" - ask people around you.)*


***


All these commands we have been using, like `summary( )` and `mean( )` are called *functions*.  A function can take all different kinds of input depending on what you are trying to do: datasets, vectors such as `bball$Team.Score`, etc. An important skill in **R** is figuring out for yourself how functions work.

For example, type `?boxplot` into your **R** console.  A help page will pop up telling you about this function.  Notice that under **Usage**, it says `boxplot(x, ...)`.  This tells you that you need to supply something called *x* to the function, and the rest of the input is optional.  But what is *x*?  Ah-ha!  There is a section called **Arguments**, which tells us that *x* is the vector of values you want to put in a boxplot.

Run the code below to make a boxplot of the team scores of college basketball games.

```{r, eval = TRUE}

# make boxplot of team scores
boxplot(bball$Team.Score)

```


***

### Question 3: 

a) Now check out `?hist`, a function for making histograms.  Below is basic code to make a histogram of `Team.Scores`, and also code for the same histogram but with a lot of the optional input changed.  Mess around with these inputs until you understand what each is doing.   

```{r, eval = TRUE}

# Boring histogram
hist(bball$Team.Score)

# Fancy histogram
hist(bball$Team.Score, breaks = 10, main = "I am a title", xlab = "I am an x-axis label", col = "grey", freq = TRUE)

```
Explain in your own words what `breaks` and `freq` change about the histogram.
```
From boring histogram to fancy histogram, breaks, which are the number of the bar shown on the graph, change from 10 to 5, and thus each bar worths from 10 points to 5 points; frequency is shown on the y-axis of the graph and changes from TRUE to FALSE and when freq=false the y-axis would show density of the scores of teams, which is frequency divided by total number of teams.
```

b) The optional inputs `main`, `xlab`, `ylab`, and `col` are common to most plotting functions.  Use what you learned in (a) to make a boxplot of `Team.Scores` with proper axis labels and title.

```{r, eval = TRUE}

# make boxplot of team scores
boxplot(bball$Team.Score, main="boxplot of bball$Team.Score",ylab="bball$Team.Score",col="grey")

```

c) To check if the histogram is Normal, or to help visualize its shape, we might want to overlay a Normal curve on top of the histogram.  The code below will do so - but the curve doesn't fit very well.  
```{r, eval = TRUE}

# Boring histogram
hist(bball$Team.Score, freq = FALSE)
xbar<-mean(bball$Team.Score)
sd<-sd(bball$Team.Score)

# overlay Normal Curve

curve(dnorm(x, mean=120, sd=20), 
      add = TRUE, col = "blue", lwd = 2)
```
Explain what the role is of the functions `curve( )` and `dnorm( )`.  Why did we put `add = TRUE` in the inputs?
```
curve() shows a curve on the graph with the information it was given; dnorm() generates a normal distribution with mean and standard deviation value; 'add= TRUE' add the curve to the existing histogram.
```

d) Alternatively, we can overlay a line that is a "smoothed" version of the data, as follows:

```{r, eval = TRUE}

hist(bball$Team.Score, freq=FALSE)
# overlay smoothed curve
lines(density(bball$Team.Score),
      col = "red", lwd = 2, lty = 2)

```

What is the difference between `lines( )` and `curve( )`?  When might we want to use `density( )`, and when would it be better to overlay a Normal curve on a histogram?
```
lines() draws by connecting line segments of the density at each points whereas curve() draws by plotting every dot on the curve. 'density()' is needed when we don't know whether the dataset is in normal distribution.It is better to overlay a Normal curve when we know the mean and standard deviation of the datasets that is in normal distribution.
```



e)  Now make your own histogram with well-chosen inputs and with a Normal overlay that fits better.  Would you say the data looks Normal?

```{r, eval = FALSE}
hist(bball$Team.Score, freq = FALSE)
xbar<-mean(bball$Team.Score)
SD<-sd(bball$Team.Score)
curve(dnorm(x, mean=xbar, sd=SD), 
      add = TRUE, col = "blue", lwd = 2)
```

***

## Subsetting

One of the most powerful qualities of **R** is the ability to select a subset of a dataset. Suppose we want to look only at games involving UNC or Duke.  We would need to figure out which rows of `bball` involve one of those teams, and then make a new dataset out of only those rows.  

For this, we will use *booleans*, which are variables with the value `TRUE` or `FALSE`.  Play around with the following code until you feel comfortable with `==`, `>`, `<`, and `%in%` as well as `&` (and) and `|` (or).

```{r, eval = TRUE}
# booleans practice

1 == 1
1 == 2
1 < 2

1 == 1 | 1  > 2
1 == 1 & 1 > 2

```

You can make up your own vector using the function `c( )`, which stands for "concatenate".  This is like making a new variable - the variable can contain anything you want, such as numbers, strings, booleans. Try the example below to make a vector and subset it. Note that we can use either `<-` or `=` to store information in a variable.

```{r, eval = TRUE}

vec <- c("cat", "dog", "horned toad", "Her Majesty Queen Elizabeth", "dog")
vec

# Some more booleans
vec == "dog"
"dog" == vec
vec %in% c("dog", "cat")
c("dog", "cat") %in% vec


# Finding indices

which(vec == "dog")
which(vec %in% c("dog", "cat"))
which(c("dog", "cat") %in% vec)

# Subsetting
new = vec[vec %in% c("dog", "cat")]
new

```

***

### Question 4:

a) The following code will give you an error.  What happened?
```{r, eval = FALSE}

vec = c(1, 2, 3, "4")
vec + 2

```
```
vec is a list of string but not a list of number; number could not be added to a string, thus "+2" is not a valid command.
```

b) The following code will NOT give you an error?  What is going on here?
```{r, eval = TRUE}

vec = c(TRUE, FALSE, FALSE, TRUE)
vec + 2

```
```
TRUE represents number 1 and FALSe represents number 0 in the coding as what true and false is presented in the binary. Thus number 2 was added to "vec", which is a list of number.
```

c) Now we are ready to make a new dataset.  We'll get a list of booleans to tell us where UNC or Duke's games are, and use that to subset the datset `bball`.

Try running each of the following lines of code.  None of them will make the datset we want.  What was the problem with each one?

```{r, eval = FALSE}

# Make new dataset with only UNC or Duke games


#A
my_subset = bball[Team == "North Carolina" | Team == "Duke", ]

#B
my_subset = bball[bball$Team == "North Carolina", bball$Team == "Duke"]

#C
my_subset = bball[bball$Team = "North Carolina" | bball$Team = "Duke", ]

#D
my_subset = bball[bball$Team == "North Carolina" & bball$Team == "Duke", ]

#E
unc_games = which(bball$Team == "North Carolina")
my_subset = bball[unc_games | bball$Team == "Duke", ]

#F
my_subset = bball[bball$Team == "North Carolina" | bball$Team == "Duke"]

```
```
#A "Team"is not the right way to retieve bball team's list; #B There should be "or" command(which is |) but not comma between two evaluantion functions; #C There should not be "==" instead of "=" as it is evaluation function but not assigning the string; #D There should be "|" instead of "&" between those evaluations; #E is wrong because for "unc_gmaes" there is no test going on, thus it is invalid to have a command like "a index OR a boolean"; #F A comma is needed behind "Duke".
```


d) Now write your own code to make the correct dataset.

```{r, eval = TRUE}
my_subset = bball[bball$Team == "North Carolina" | bball$Team == "Duke",]
my_subset

```

***

## Z-Scores and t-scores

Alright, enough of that data wrangling.  Time to do some statistics.

Check out `?Normal`.  These are some functions that will help us calculate probabilities about the Normal distribution. (No more using Table A!)  The most important ones are `pnorm` and `qnorm`.

`pnorm(q)` will tell you the probability of a standard Normal being below the value `q`

`qnorm(p)` will tell you the z-score that has area `p` below it on a standard Normal curve

***

### Question 5

a) For each of the following lines of code, think about what the result will be **before** running the code.  **Draw a picture for each one to visualize what is going on with `pnorm` and `qnorm`.**

```{r, eval = FALSE}
# practice with Normal densities in R

#i
pnorm(0)
qnorm(0)

#ii
pnorm(100)
qnorm(100)

#iii
qnorm(pnorm(0))
qnorm(pnorm(7))

#iv
pnorm(qnorm(0))
pnorm(qnorm(0.5))

#v
pnorm(0, sd = 10)
pnorm(0, mean = 1, sd = 10)

#vi
qnorm(0.05)
qnorm(0.05, sd = 10)
qnorm(0.05, mean = 1, sd = 10)

```

b) Why did you get an error in part (ii)?
```
```

***

Now use this code to make a new variable for the total score of a game:

```{r, eval = FALSE}
# Make new variable
bball$Total.Score = bball$Team.Score + bball$Opponent.Score
```

We will use *z-scores* and *t-scores* to think about whether a game is unusually high scoring.

***

### Question 6:

a) As you may have noticed, the dataset `bball` actually displays each game twice: once for each team.  Make a new dataset with each game listed only once by subsetting `bball`.
```{r, eval = FALSE}
YOUR CODE HERE
```

b) On Feb 17, 2016, UNC played Duke.  Using the Normal distribution, what percent of games have higher scores than the UNC/Duke game?  (Assume that the mean and standard deviation of `Team.Score` are actually the *population* mean and standard deviation.)  
```{r, eval = FALSE}
YOUR CODE HERE
```

c) What percentage of games in the dataset did we observe to be higher scoring than the UNC/Duke game?  The functions `sum( )` and `length( )` will help you answer this question.

```{r, eval = FALSE}
YOUR CODE HERE
```

d)  What is the difference between what we did in (b) and (c)?  Do you think the Normal approximation is reasonable for this data?  Why or why not?
```
```

***

Recall that *t-scores* are used instead of *z-scores* when the population standard deviation is unknown.  The functions `pt` and `qt` work almost same way as `pnorm` and `qnorm`, but for the t-distribution instead of the Normal.  However, be careful, and read `?pt` for help!  These functions don't let you enter the mean and standard deviation as input - you need to figure out what do about that!

***

### Question 7:
Use all your new **R** skills to answer this question: Was the Feb 17th game between UNC and Duke particularly high scoring *for a UNC game*?

```{r, eval = FALSE}
YOUR CODE HERE

```

***

## Confidence Intervals and Proportions

You now have all the **R** knowledge you need to make some confidence intervals!  You may wish to go over your lecture notes for this section, especially to remind yourself how to deal with proportions.

***

### Question 8:

a) Make a 95% confidence interval for the number of points UNC scores in a given game.  You will need to think about which **R** commands will give you critical values of the *t*-distribution, and how to use these to make a confidence interval.

```{r, eval = FALSE}
YOUR CODE HERE

```

b) What percentage of games did UNC win in 2015-2016?  Make a 95% confidence interval for their win percentage.


```{r, eval = FALSE}
YOUR CODE HERE
```

***

## Hypothesis Testing

You have now had lots of practice learning to use a function by reading the documentation.  Part of the point of this course is for you to become familiar enough with **R** to learn new commands and functions without being shown how to use them.  This will make you a skillful (and hireable!) programmer in the future.

Check out `?t.test` and `?prop.test`.  Figure out what these functions do, what input they take, etc. Then answer the following questions.

***

### Question 9:

a) Does UNC tend to win more games than they lose?  That is, is there evidence at the 0.05 level that the "true" probability of UNC winning a given game in 2015-2016 is larger than 0.5?

```{r, eval = FALSE}
YOUR CODE HERE

```

b) Based on how many points they tend to score in a game, would you say UNC and UCLA were equally good teams? 

```{r, eval = FALSE}

YOUR CODE HERE

```

c) Based on win percentage, would you say UNC and UCLA were equally good teams? 

```{r, eval = FALSE}
YOUR CODE HERE

```

***

## Do your own analysis

Now it's your turn to think like a Statistician.  

Many sports fans believe that teams tend to play better when they are in their home arena, perhaps because they are more comfortable there or because they are energized by their fans.  This idea is called "Home Court Advantage".

Based on the `bball` data, do you think there is any evidence of Home Court Advantage?  Be creative - think about different ways you might measure if a team is playing "better" at home.

Explain and justify your answer.